### 2024-02-08 doc/15yrsChange.Rmd
Add Outline of Recruitment Failure

*Summary formats tend to be in a few long bullet points

@@

@@

The study was conducted at The Smithsonian Conservation Biology Institute’s (SCBI) large forest dynamics plot in Front Royal, Virginia, USA (38.8935°, -78.1454°; Bourg et al. 2013, Anderson-Teixeira et al. 2015). *A sentence about the region and that it is adjacent to the northern entrance of Shenandoah National Park.* The plot is 25.6 ha (640 x 400 m) and includes a 4-ha, 2.4 m high deer exclusion fence, which has decreased the number/decreased the presence of white-tailed deer since 1990 (McShea 2000). It consists of a mature secondary eastern deciduous forest that developed following logging and agricultural use in the mid-19th century. The species composition reflects what is typical of this forest type, with a canopy that is dominated by tulip poplar (*Liriodendron tulipifera*), oak (*Quercus spp.*), and hickory (*Carya spp.*), and the understory that is primarily composed of spicebush (*Lindera benzoin*), paw-paw (*Asimina triloba*), American hornbeam (*Carpinus caroliniana*), and witch hazel (*Hamamelis virginiana*)

>>>>

*Summary format for New Phytologist tends to be in a few long bullet points*

@@

Recruitment Failure Paragraph Outline

- Nuisance species have a profound impact on regional ecosystems
- There is a documented overabundance of white-tailed deer (*Odocoileus virginianus*) (*Publication referenced in McGravey et al. 2013*) which can be attributed to human influence (*Publication referenced in McGravey et al 2013*)
- White-tailed deer consume seedlings and saplings, negatively impacting seedling and sapling survival and density (*Publications referenced in McGravey et al. 2013*)
- Define or contextualize non-endemic plants (Look for a reference)
  - Working definition: Plants that were not historically present in an area, region or ecosystem or were found at different abundances or densities
- Non-endemic seedlings and saplings may not be consumed at similar rates as endemic plants (alluded to in *McGravey et al. 2013*) and may be able to utilize niche space in a forest faster or more efficiently (Reference)
- Additional information on the effects of non-endemic species
- As a result, they may be able to out compete endemic species
- These nuisance species can contribute to recruitment failure
  - Consider defining and contextualizing recruitment failure a little more here
  - The effects of recruitment debt on mature tree species composition and forest structure are often not apparent for decades (*McGravey et al. 2013* - potentially move to regeneration debt paragraph)

@@

*What is the best way to describe the forest type in our study site? I've seen "mixed mature deciduous forest", "mature secondary eastern deciduous forest", and "mesic temperate deciduous forest".  I want to make sure I'm being consistent.*

The study was conducted at The Smithsonian Conservation Biology Institute’s (SCBI) large forest dynamics plot in Front Royal, Virginia, USA (38.8935°, -78.1454°; Bourg et al. 2013, Anderson-Teixeira et al. 2015). *A sentence about the region and that it is adjacent to the northern entrance of Shenandoah National Park.* The plot is 25.6 ha (640 x 400 m) and includes a 4-ha, 2.4 m high deer exclusion fence, which has decreased the number/decreased the presence of white-tailed deer since 1990 (McShea 2000). It consists of a mature secondary eastern deciduous forest that developed following logging and agricultural use in the mid-19th century. *Canopy trees are 84-124 years old (Thompson and McGarvey - unpublished data referenced in McGarvey 2013 - see if you can find the published version of this data)* The species composition reflects what is typical of this forest type, with a canopy that is dominated by tulip poplar (*Liriodendron tulipifera*), oak (*Quercus spp.*), and hickory (*Carya spp.*), and the understory that is primarily composed of spicebush (*Lindera benzoin*), paw-paw (*Asimina triloba*), American hornbeam (*Carpinus caroliniana*), and witch hazel (*Hamamelis virginiana*)
